---
title: "Untitled"
author: "JFynn"
date: "11 February 2015"
output: html_document
---

# Loading Data and Packages

```{r}

library("ggplot2")
library("reshape2")
library("dplyr")
library("corrplot")
library("class")
library("RColorBrewer")
library("randomForest")


setwd('/Users/Flynn/Documents/projects/titanic_kaggle')

train <- read.csv('./data/train.csv', stringsAsFactors=FALSE)
test <- read.csv('./data/test.csv', stringsAsFactors=FALSE)
gendermodel <- read.csv('./data/gendermodel.csv', stringsAsFactors=FALSE)
genderclassmodel <- read.csv('./data/genderclassmodel.csv', stringsAsFactors=FALSE)

```


# Style Guide

```{r}

titanic_theme <- function() {

  # Generate the colors for the chart procedurally with RColorBrewer
  palette <- brewer.pal("Greys", n=9)
  color.background = palette[2]
  color.grid.major = palette[3]
  color.axis.text = palette[6]
  color.axis.title = palette[7]
  color.title = palette[9]

  # Begin construction of chart
  theme_bw(base_size=9) +

  # Set the entire chart region to a light gray color
  theme(panel.background=element_rect(fill=color.background, color=color.background)) +
  theme(plot.background=element_rect(fill=color.background, color=color.background)) +
  theme(panel.border=element_rect(color=color.background)) +

  # Format the grid
  theme(panel.grid.major=element_line(color=color.grid.major,size=.25)) +
  theme(panel.grid.minor=element_blank()) +
  theme(axis.ticks=element_blank()) +

  # Format the legend, but hide by default
  theme(legend.background = element_rect(fill=color.background)) +
  theme(legend.text = element_text(size=7,color=color.axis.title)) +

  # Set title and axis labels, and format these and tick marks
  theme(plot.title=element_text(color=color.title, size=10, vjust=1.25)) +
  theme(axis.text.x=element_text(size=7,color=color.axis.text, angle = -45, 
                                 vjust = 1.05, hjust =0 )) +
  theme(axis.text.y=element_text(size=7,color=color.axis.text)) +
  theme(axis.title.x=element_text(size=12,color=color.axis.title, vjust=0)) +
  theme(axis.title.y=element_text(size=12,color=color.axis.title, vjust=1.25))

}


```




# Basic Plotting

```{r}

ggplot(train, aes(Age, Fare, color = Survived)) +
        geom_point() +
        titanic_theme()


ggplot(train, aes(Embarked, fill = ..count..)) +
        geom_bar() +
        theme_bw()


```

# Cleaning


```{r}

# Change 'Embarked' values to something more easily
# understood 

train$Embarked <- as.character(train$Embarked)

train$Embarked[train$Embarked == 'C'] <- 'Cherbourg'
train$Embarked[train$Embarked == 'Q'] <- 'Queenstown'
train$Embarked[train$Embarked == 'S'] <- 'Southampton'
train$Embarked[train$Embarked == ' '] <- 'Unknown'

test$Embarked[test$Embarked == 'C'] <- 'Cherbourg'
test$Embarked[test$Embarked == 'Q'] <- 'Queenstown'
test$Embarked[test$Embarked == 'S'] <- 'Southampton'
test$Embarked[test$Embarked == ' '] <- 'Unknown'


# Create Age Groups based on the 
train$Age_bin <- cut(train$Age, breaks = seq(0, 100, by = 10))
train$Age_bin <- as.character(train$Age_bin)

train$Age_bin[train$Age_bin == '(0,10]'] <- 'Child'
train$Age_bin[train$Age_bin == '(10,20]'] <- 'Teen'
train$Age_bin[train$Age_bin == '(20,30]'] <- '20s'
train$Age_bin[train$Age_bin == '(30,40]'] <- '30s'
train$Age_bin[train$Age_bin == '(40,50]'] <- '40s'
train$Age_bin[train$Age_bin == '(50,60]'] <- '50s'
train$Age_bin[train$Age_bin == '(60,70]'] <- '60s'
train$Age_bin[train$Age_bin == '(70,80]'] <- 'Old'
train$Age_bin[train$Age_bin == '(80,90]'] <- 'Really Old'
train$Age_bin[train$Age_bin == '(90,100]'] <- 'Probably Too Old'


test$Age_bin <- cut(test$Age, breaks = seq(0, 100, by = 10))
test$Age_bin <- as.character(test$Age_bin)

test$Age_bin[test$Age_bin == '(0,10]'] <- 'Child'
test$Age_bin[test$Age_bin == '(10,20]'] <- 'Teen'
test$Age_bin[test$Age_bin == '(20,30]'] <- '20s'
test$Age_bin[test$Age_bin == '(30,40]'] <- '30s'
test$Age_bin[test$Age_bin == '(40,50]'] <- '40s'
test$Age_bin[test$Age_bin == '(50,60]'] <- '50s'
test$Age_bin[test$Age_bin == '(60,70]'] <- '60s'
test$Age_bin[test$Age_bin == '(70,80]'] <- 'Old'
test$Age_bin[test$Age_bin == '(80,90]'] <- 'Really Old'
test$Age_bin[test$Age_bin == '(90,100]'] <- 'Probably Too Old'

```



## Data Aggregation

This is an attempt to write a reproducible way to get very detailed summary statistics for Titanic data.

This for first chunk will tell you the survival percentage of people who embarked from Southampton

```{r}

summary_stats <- group_by(train, Embarked, Survived)
summary_stats <- summarise(summary_stats, 
                           count = n()
                           ) 

summary_stats$Survived[summary_stats$Survived == 0] <- 'dead'
summary_stats$Survived[summary_stats$Survived == 1] <- 'survived'

summary_stats <- dcast(summary_stats, Embarked ~ Survived)
summary_stats[is.na(summary_stats)] <- 0

summary_stats <- mutate(summary_stats,
                         non_survival_rate = summary_stats$survived / 
        (summary_stats$dead + summary_stats$survived))



summary_stats


```

This chunk will tell you the survival rates of passengers depending on how old they are, their gender, and the number of siblings which they have on the boat. 


```{r}



summary_stats <- group_by(train, Sex, Age_bin, SibSp, Survived)
summary_stats <- summarise(summary_stats, 
                           count = n()
                           ) 

summary_stats$Survived[summary_stats$Survived == 0] <- 'dead'
summary_stats$Survived[summary_stats$Survived == 1] <- 'survived'

summary_stats <- dcast(summary_stats, Sex + Age_bin + SibSp ~ Survived)
summary_stats[is.na(summary_stats)] <- 0

summary_stats <- mutate(summary_stats,
                         non_survival_rate = summary_stats$survived / 
        (summary_stats$dead + summary_stats$survived))



summary_stats

```



# Random forest

```{r}

train_ed <- train[-4]

train_ed <- train_ed[-8]
train_ed <- train_ed[-9]
train_ed <- train_ed[-10]
train_ed <- train_ed[-1]

train_ed$Sex <- as.factor(train_ed$Sex)
train_ed$Embarked <- as.factor(train_ed$Embarked)
train_ed$Survived <- as.factor(train_ed$Survived)


train.imputed <- rfImpute(Survived ~ ., train_ed)

train.rf <- randomForest(Survived ~ ., data=train.imputed, importance=TRUE,
                         proximity=TRUE)


test_ed <- test[-3]
test_ed <- test_ed[-1]
test_ed <- test_ed[-6]
test_ed <- test_ed[-7]
test_ed <- test_ed[-8]
test_ed$Sex <- as.factor(test_ed$Sex)
test_ed$Embarked <- as.factor(test_ed$Embarked)



# train.pred <- predict(train.rf, test_ed, type = 'response')

```



```{r}



# titanic_class <- knn(train, test, cl = train_labesl, k = 29)


```

